---
title: "Which Book To Read?"
author: "Sarah Heidmann"
output: html_document
---
*Data through 1 Apr 2025*
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(tidyverse)
library(lubridate)
library(knitr)
books <- read_csv("data/goodreads_library_export_20250401.csv",
                  col_types = cols(`Date Read`=col_date(format="%Y/%m/%d"),
                                   `Date Added`=col_date(format="%Y/%m/%d")))
widebooks <- books %>%
      separate_rows("Bookshelves", sep = ", ") %>%
      select(Bookshelves, Title) %>%
      mutate(shelflog = 1) %>%
      pivot_wider(names_from=Bookshelves,
                  values_from=shelflog) %>%
      mutate_all(replace_na,0)

wanttoread <- widebooks %>% 
   filter(`to-read`==1) %>%
   nrow()
accessible <- widebooks %>% 
   filter(`to-read`==1) %>%
   filter(kindle==1|library==1|`st-thomas`==1|audible==1|`library-open`==1) %>%
   nrow()
fiction <- widebooks %>% 
      filter(`to-read`==1 & fiction==1) %>% 
      nrow()
nonfiction <- widebooks %>% 
      filter(`to-read`==1 & nonfiction==1) %>% 
      nrow()

booksperday_365 <- books %>%
   filter(`Date Read`>=(today()-365)) %>%
   nrow() / 365
rawdaysleft <- wanttoread / booksperday_365
booksaddedperday_365 <- books %>%
   filter(`Date Added`>=(today()-365)) %>%
   nrow() / 365
netdaysleft <- wanttoread / (booksperday_365 - booksaddedperday_365)

booksperday_30 <- books %>%
   filter(`Date Read`>=(today()-30)) %>%
   nrow() / 30
booksaddedperday_30 <- books %>%
   filter(`Date Added`>=(today()-30)) %>%
   nrow() / 30
```

There are `r wanttoread` books on my Want-To-Read shelf.
I have access to `r accessible` of them. Of my TBR, `r fiction` are fiction and `r nonfiction` are nonfiction.

In the past year, on average, I read a book every `r round(1/booksperday_365,1)` days.
If I continued to read at this rate and added no more books, it would take me `r round(rawdaysleft,0)` days (`r round(rawdaysleft/365,1)` years) to finish reading the rest of my list.

In the same time, I added a book every `r round(1/booksaddedperday_365,1)` days.
If I continued to add and read at this rate, it would take me `r round(netdaysleft,0)` days (`r round(netdaysleft/365,1)` years) to finish reading the rest of my list.

In the past month, I read a book every `r round(1/booksperday_30,1)` days, and added a book every `r round(1/booksaddedperday_30,1)` days.


## Highest Rated Books

The 10 highest rated unread books in my library that I already have access to:

```{r highRating}
books %>%
   filter(`Exclusive Shelf`=="to-read") %>%
   select(Title, Author, `Original Publication Year`,
          `Number of Pages`,`Average Rating`) %>%
   left_join(select(widebooks,Title,
                    kindle,library,`st-thomas`,audible,`library-open`), 
             by="Title") %>%
   filter(kindle==1|library==1|`st-thomas`==1|audible==1|`library-open`==1) %>%
   pivot_longer(cols=all_of(c("kindle","library","st-thomas",
                              "audible","library-open")),
                names_to = "Available") %>%
   filter(value==1) %>%
   select(-value) %>%
   arrange(desc(`Average Rating`)) %>%
   slice_head(n=10) %>%
   kable()
```

## Rankings

Unread books currently available to me either in person, on my Kindle, or through Libby, ranked.
Book ranking is the number of satisfied conditions below:

* average rating above 4.0
* highly recommended
* island-themed or related
* recommended by Emily Graslie or my teacher Mike
* on a "life-changing" list
* on a "top 100 lifetime" list
* won an award
* made into a movie

The result, my top-10 to-read books:
```{r highRanking}
widebooks %>%
   left_join(select(books, Title, Author, `Original Publication Year`,
                    `Number of Pages`, `Average Rating`), 
             by="Title") %>%
   mutate(highrating = ifelse(`Average Rating`>4,1,0)) %>%
   filter(`to-read`==1) %>%
   filter(kindle==1|library==1|`st-thomas`==1|audible==1|`library-open`==1) %>%
   pivot_longer(cols=all_of(c("kindle","library","st-thomas",
                              "audible","library-open")),
                names_to = "Available") %>%
   filter(value==1) %>%
   mutate(ranking=(`highly-recommended`*2)+island+`life-changing`+ highrating +
             `top-100-lifetime`+`award-winning`+`brain-scoop`+`from-mike`+
             `made-into-movie`) %>%
   select(Title, Author,`Original Publication Year`, ranking, 
          `Number of Pages`, `Average Rating`, Available) %>%
   arrange(desc(ranking), desc(`Average Rating`)) %>%
   slice_head(n=10) %>%
   kable()
```

