---
title: "To Find"
author: "Sarah Heidmann"
output: html_document
---
*Data through 29 Apr 2023*
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(tidyverse)
library(knitr)
books <- read_csv("data/goodreads_library_export_20230429.csv",
                  col_types = cols(`Date Read`=col_date(format="%Y/%m/%d"),
                                   `Date Added`=col_date(format="%Y/%m/%d")))
widebooks <- books %>%
      separate_rows("Bookshelves", sep = ", ") %>%
      select(Bookshelves, Title) %>%
      mutate(shelflog = 1) %>%
      pivot_wider(names_from=Bookshelves,
                  values_from=shelflog) %>%
      mutate_all(replace_na,0)
```

## Reverse rankings

Much of my library was added in 2014, assembled with little thought from very old to-read lists I have been keeping my whole life.

To help slowly chip away at books on my list that no longer pique my interest, here is a reverse ranking of books to look at for deletion or better classification:
```{r lowRanking}
widebooks %>%
      left_join(select(books,Title,`Average Rating`), by="Title") %>%
      filter(`to-read`==1) %>%
      filter(kindle==0 & library==0 & `st-thomas`==0 & 
                   audible==0 & `library-open`==0 & 
                   `Average Rating`<4) %>%
      mutate(shelves=rowSums(select(.,-Title,-`Average Rating`))) %>%
      mutate(ranking=(`highly-recommended`*2)+island+`life-changing` +
             `top-100-lifetime`+`award-winning`+`brain-scoop`+`from-mike`+
             `made-into-movie`) %>%
      select(Title, ranking, shelves,`Average Rating`) %>%
      arrange(ranking,shelves,`Average Rating`) %>%
      slice_head(n=20) %>%
      kable()
```

## Books to look for

Now let's reverse that. There are many books I don't have that may be worth looking harder for or buying.

Here is a ranking of books to find a way to get access to:
```{r toFind}
widebooks %>%
      left_join(select(books,Title,`Average Rating`), by="Title") %>%
      filter(`to-read`==1) %>%
      filter(kindle==0 & library==0 & `st-thomas`==0 & 
                   audible==0 & `library-open`==0 &
                   `Average Rating`>4) %>%
      mutate(shelves=rowSums(select(.,-Title,-`Average Rating`))) %>%
      mutate(ranking=(`highly-recommended`*2)+island+`life-changing` +
             `top-100-lifetime`+`award-winning`+`brain-scoop`+`from-mike`+
             `made-into-movie`) %>%
      select(Title, ranking, shelves,`Average Rating`) %>%
      arrange(desc(ranking),shelves,`Average Rating`) %>%
      slice_head(n=20) %>%
      kable()
```

